Vagueness, Noise, and Communication

### 2014-08-11 paper/03_simulations.tex
Generated new plots, included them in the results section.
Added short description of results. Renamed results section to discussion.

\subsection{Measures}
To be able to quantitively analyze the results of our simulations, we define a number of measures to capture different characteristics of sim-max languages.

@@

We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
For each parameter combination, we perform $100$ trials.

>>>>

\subsection{Measures}
\label{sec:measures}
To be able to quantitively analyze the results of our simulations, we define a number of measures to capture different characteristics of sim-max languages.

@@

We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
For each parameter combination, we perform $100$ trials and record the values of each of the measures introduced in Section~\ref{sec:measures} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.

\subsection{Results}
In order to visualize the impact of the impairment factor on the outcome of simulations, in Figure~\ref{fig:metrics-boxplots} we plot the distribution of each of the measurements per impairment value.
%
\begin{figure}
        \centering
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140121-141158}
                \caption{Entropy of sender strategy.}
%                \label{fig:gull}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140121-141158}
                \caption{Entropy of receiver strategy.}
        \end{subfigure}

        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140121-141158}
                \caption{Informativity of sender strategy.}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140121-141158}
                \caption{Informativity of receiver strategy.}
        \end{subfigure}

        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140121-141158}
                \caption{Voronoiness of sender strategy.}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140121-141158}
                \caption{Voronoiness of receiver strategy.}
        \end{subfigure}

        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Expected-utility-20140121-141158}
                \caption{Expected utility.}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \includegraphics[width=\textwidth]{plots/Iterations-20140121-141158}
                \caption{Number of iterations until convergence.}
        \end{subfigure}
        \caption{Boxplots of metrics per impairment value.}
        \label{fig:metrics-boxplots}
\end{figure}
%
Most trends are quite clear.
Entropy seems to increase with higher impairment values for both sender and receiver.
This is an expected outcome, since the higher the impairment value, the more likely each player is to confuse two states, thus the more uncertain he will be in unequivocally associating a state with a message, or vice-versa.
The trend for informativity is also expected: more vague strategies convey less information, be it from the sender or the receiver perspective.

Regarding Voronoiness, we observe that most simulation results approximate perfect Voronoi tessalations.
This is not the case, however, for the lower values of impairment.
We believe the explanation is that, with lower values of impairment, strategies can achieve a high expected utility without having perfect Voronoiness.
Although the expected long term outcome in these situations is perfect Voronoiness (given J\"ager's analytical results~\cite{Jager2007}), actually reaching it involves minor refinements which at some point do not qualify as significant change given our convergence criterion.

The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009}: more impairment necessarily leads to lower expected utility.
Finally, the plot for number of iterations until convergence shows that a certain level of impairment enables faster convergence.
We will discuss this further in more detail in Section~\ref{sec:discussion}.

\mytodo{JPC}{Present and discuss results per size of state space? Don't forget O'Connor's claim that speed of convergence is more important for larger \#state/\#message ratio}

### 2014-08-11 paper/04_discussion.tex
Added short description of results. Renamed results section to discussion.

\appendix

\section{Proofs}
\label{sec:proofs}

\begin{proof}[Proof of Theorem~\ref{thm:sender-eq}]
  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
  the consequent:
  \begin{align*}
    C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
    \cdot \Sstrat(\messg_y \probbar \state_l) \\
    & =  \sum_{\state_l} C_{xl}
    \cdot  \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \\
    & = \sum_{\state_l}
    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
  \end{align*}
  Next consider the lhs of the consequent:
  \begin{align*}
    G(Q^C(\Smixed))(\messg_y \probbar \state_x) & =
    \sum_{\Spure_i(\state_x)=\messg_y} Q^C(\Smixed)(\Spure_i) \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed(\Spure_j) \cdot Q^C_{ji} \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed(\Spure_j) \cdot \prod_{\state_l} \sum_{\state_m \in
      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} \\
    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
  \end{align*}
  To simplify this further we look at a fixed $\Spure_j$ and consider
  the term: 
  \begin{align}
    \label{eq:term}
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
  \end{align}
  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
  over all pure strategies, we essentially look at each such
  mapping. Term (\ref{eq:term}) above quantifies over all pure
  strategies that map $\state_k$ onto $\messg_y$. The above term then
  sums over all products whose factors are tuples in $\times_{k>2}
  \set{y \setbar \exists l \mycolon y = Y_{kl}}$. So term
  (\ref{eq:term}) expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
  \begin{align*}
    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
    \dots \\
    & + (Y_{11} \cdot Y_{2d} \cdot
    Y_{3d} \cdot \ldots \cdot
    Y_{ed})\\
    = &    
  \end{align*}
  But since $Y$ is a row-stochastic matrix, this simplifies to
  $Y_{xy}$. Continuing the derivation with this:
  \begin{align*}
    G(Q^C(\Smixed))(\messg_y \probbar \state_x) 
    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed(\Spure_j) \cdot
     C_{xl} \\
     & = \sum_{\state_l}
    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
  \end{align*}

\end{proof}

\begin{proof}[Proof of \ref{thm:receiver-eq} ]
  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
  expands to:
  \begin{align*}
    C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
    \Rmixed_i \cdot C_{jx} \\
    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
    \Rmixed_i \cdot C_{jx} \\
    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
  \end{align*}
  The rhs expands to:
  \begin{align*}
    G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{\Rpure_i}
    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
        \state_x}} 
    \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}
  \end{align*}
  Without loss of generality, assume that $x=y=1$, and fix
  $\card{M}=d$ and let $e$ be the number of pure receiver
  strategies. Then the last term can be rewritten as:
  \begin{align*}
    & = \sum_{i}
    \Rmixed_i \cdot ( C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_1(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_1(\messg_d)} + \ldots  \\
      & \textcolor{white}{ = \sum_{i}
    \Rmixed_i  (}  +  C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_2(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_2(\messg_d)} + \ldots  \\
      & \textcolor{white}{ = \sum_{i}
    \Rmixed_i  (}  + C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_e(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_e(\messg_d)}) )
  \end{align*}
  Every for every messages $C_{\Rpure_i(\messg_l)}$ is a stochastic
  vector. For $l>1$, all elements of these vectors appear equally
  often. But that means that these cancel out. What remains is:
  \begin{align*}
    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
  \end{align*}
\end{proof}

>>>>

\section{Discussion}
\label{sec:discussion}

### 2014-08-11 paper/paper.tex
paper skeleton first content
Added description of first metric: entropy.
Generated new plots, included them in the results section.
Added short description of results. Renamed results section to discussion.

\usepackage{myarticlestyledefault}

\title{Vagueness, Noise and Communication}
\author{Jos\'e Pedro Correia and Michael Franke}
\date{}

@@

@@

\input{04_results}

@@

>>>>

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
\bibliography{../helpers/MyRefGlobal}
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\title{Vagueness, Noise, and Communication}
\author{Jos\'e Pedro Correia and Michael Franke}
\date{}

@@

\begin{abstract}
\mytodo{JPC}{This start needs to be adjusted depending on the audience}
Signaling games have been productively used as models for studying natural language meaning.
One important aspect that is, however, typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
Recently, some models have been proposed to explicitly address this limitation.
In this paper we explore one such models.
Namely, we establish a link between the dynamics in the model and the replicator-mutator equations.
Furthermore, we study the long term behavior of the model using computer simulations.
\mytodo{JPC}{Kind of a big claim below\ldots}
Our results indicate that a possible explanation of why vagueness is pervasive in natural language is that vague languages converge faster than non-vague ones, thus providing an evolutionary advantage in an changing environment.
\end{abstract}

@@

\input{04_discussion}

@@

\input{06_appendix}
